---
title: "Tutorial"
author: "Josh Rivera"
date: "11/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Howdy! The first thing one should do to utilize this set of functions is install the Rivera package. *code seen below*

```{r, eval = FALSE}
devtools::install_github("joshoriv/rivera.package")
```

Once the package has been installed, please be sure to load the package into your R console.

```{r}
library(rivera.package)
```

At this point, we can go ahead and load the required packages for this tutorial. Following this we'll be reading in the siren locality data and siren shape data (landmark coordinates).

```{r}
library(readr)
library(ggplot2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(mapdata)
library(maps)
library(rgl)
library(geomorph)
library(svd)
library(shapes)
sirens <- readr::read_csv("../inst/extdata/sirens_updated.csv")
file_path <- system.file("extdata", "SirenDorsal.TPS", package = "rivera.package")
sirens_dorsal <- geomorph::readland.tps(file_path, specID = "ID")

  
```

## Specimen Locality Distribution Map

The purpose of this function is to generate a map of siren specimen localities. Specimen distribution maps are helpful in understanding trends in biological phenomena. In order for this function to process, the coordinate information (longitude & latitude) must be available for each specimen. The world map, where the specimen datapoints will be superimposed, will be selected by providing coordinates of interest.

```{r}

sirens_map <- dplyr::select(sirens, Longitude, Latitude) %>% 
  na.omit()
world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) +
  geom_sf() +
  geom_point(data = sirens_map, mapping = aes(x = Longitude, y = Latitude), color = "blue") +
  coord_sf(xlim = c(-110, -75), ylim = c(22, 50), expand = FALSE)
  
```

## Generalized Procrustes Analysis

This function performs a generalized Procrustes analysis/superisposition of multiple specimens about their mean. This function requires the input of specimen landmark coordinates. A plot of superimposed coordinates is returned and their centroid sizes in dataframe formate. This function then plots the Procrustres coordinates. This is a useful analysis for visualizing subtle variation in morphology.

```{r}

gpa <- gpagen(sirens_dorsal)
plot(gpa)
  
```

## Principal Components Analysis

This function performs a traditional principal components analysis. The input for this function is the generalized Procrustes analysis proceeding this function. A PCA plot shows clusters of samples based on their similarity. This can be helpful to visualize similarities among geographically isolated taxa, different species, etc.

```{r}

pca <- gm.prcomp(gpa$coords)
  summary(pca)
  plot(pca, main = "PCA")
  
```
